---
title: "Seminar 5 Solved xdxd"
author: "Paolo Ginatta"
date: "3/20/2019"
output:
  html_document: default
  pdf_document: default
  
---
<style>
p.caption {
  font-size: 0.9em;
  font-style: italic;
  color: grey;
  margin-right: 10%;
  margin-left: 10%;  
  text-align: center;
}
</style>


```{r,echo=FALSE,warning=FALSE,}
library(knitr)
knit_hooks$set(small.mar = function(before, options, envir) {
    if (before) par(mar = c(4, 5, .3, 4))  
})
```
            


---


##### First Excersice
<div style="text-align: justify"> 
Both the confidence interval and the hypothesis test are found in the prop.test. `alternative = "greater"` was used because the problem suggests that men have a _greater_ probability of having the disease.
</div>

```{r, eval=TRUE}
prop.test(59,100, conf.level=0.90,alternative = "greater") 
```
<div style="text-align: justify"> 
The null probability **0.5** is trapped in the interval **[0.5209788 1.0000000]** of the sample estimate p, that is **0.59**. The probability that the test statistic (in this case a X-squared with 1 degree of freedom ) is of the value obtained (2.89) or higher, is that of the _p-value_, that is **0.04457**. With this, the null hypothesis that the probability was _lower or equal_ to 0.5 _can_ be rejected.
</div>

\newpage

##### Second Excersice
<div style="text-align: justify"> 
In this problem a _paired sample_ is presented with a _before_ and an _after_ levels of cholesterol, and it is required to test if the _after_ is **smaller** than the _before_ level. First, the samples are introduced.
</div>
```{r}
before <-c(200,210,330,240,260,300,245,210,190,225)
after <-c( 150,200,275,250,200,250,200,180,190,205)
```
<div style="text-align: justify"> 
Now we do a t test of the two variables with the option `paired=TRUE` enabled.
</div>
```{r,echo=TRUE}
t.test(before,after,paired = TRUE,alternative = "greater")
```

<div style="text-align: justify"> 

Note that the differences of the means are _before_ - _after_. So, if the treatment is effective the result should be positive as _after_ would be smaller, that's why `alternative = "grater"` was used, since (if effective) it would have to be _grater_ than 0. And, because the p-value is smaller than 5%, then it can be concluded that the null hypothesis _smaller or equal to 0_ can be rejected, or in other words, it can be said that the treatment _is efective_ with a **5%** significance level.
 </div>

\newpage
##### Third Excersise.
<div style="text-align: justify"> 
In this exercise it is asked to see whether the two samples' averages are actually different. First, we introduce the data.
</div>
```{r,echo=TRUE}
linea <- c(10,30,32,23,23,24,20,18,39,35)
lineb <- c(12,28,30,20,25,31,15,12,22,24,40,30)
```
Now, the samples are compared graphically.
```{r plot 1,fig.align="center",fig.cap ="Figure: Boxplot of the samples of lines A and B" }
boxplot(linea,lineb,names = c("A","B"),ylab = "passengers",xlab = "lines",
col = "seashell",cex.lab=0.80, cex.axis=0.80, cex.main=1.5, cex.sub=0.8)
```


<br><br><br>
<div style="text-align: justify"> 
It can be appreciated in the box plots that they are slightly different, but before doing a test to see if they actually are different (and not only by chance) a test to see if the variances are _equal_ is needed, that is `var.test`.
</div>
```{r}
var.test(linea,lineb)
```
<div style="text-align: justify"> 
As seen in the result of the test the ratio of the two variables is _not_ equal to 1 by a little, in fact, the probability that this inequality happens, being actually equal to 1, is of 0.91, so we _cannot_ reject the fact that they **variances** are equal.

Now that it is known that the variances are equal, a _two sample t-test_ can be done in order to see if the means are actually different. Note that `var.equal = TRUE` is used since  their equality has been already verified, and `mu = 0` since the problem asks to check only if the means are _diferent_ from each other (alternative: _not_ equal to 0).
</div>
```{r}
t.test(linea,lineb, mu=0, conf.level=0.90, var.equal=TRUE)
```
<div style="text-align: justify"> 
The **sample estimates** at the end of the result shown above indicate that the means are different, but the 90% confidence interval of this difference traps **0** (that is the null hypothesis) and a **p-value** of **0.7226** suggests that the probability of this difference happening, being the means actually _equal_, is too high for it to be taken in consideration. As a result, it can be said with a 10% significance level, that means _are_ equal.
</div>


\newpage
##### Fourth Exercise
<div style="text-align: justify"> 
In this one, the same as exercise three has to be done but now between lines _A_ and _C_. First, line C is introduced (line A was introduced in the previous exercise)
</div>
```{r}
linec <- c(32,39,35,30,37,28,34,33,25,30,37,33)
```
Now, the samples are compared graphically.

```{r plot2,fig.align="center",fig.cap ="Figure: Boxplot of the samples of lines A and C"}
boxplot(linea,linec,names = c("A","C"),ylab = "passengers",xlab = "lines",
col = "seashell",cex.lab=0.80, cex.axis=0.80, cex.main=1.5, cex.sub=0.8)
```
<br><br><br>
<div style="text-align: justify"> 
Now the difference is much more evident, and the variances may not be equal. So a `var.test` is done before testing the means' difference too see if variances are equal to make a test like before or a _Welch_ test for different variances.
</div>
```{r}
var.test(linea,linec)
```
<div style="text-align: justify"> 
From the results of the test it can be concluded that the variances are **not** equal, so a _Welch_ test will be done. For this, a _**plain** two-sample t.test_ can be done, since it uses `var.equal = FALSE` as the default option. Remember that `mu = 0` is used since the null hypothesis is that they are _equal_ (the difference between them is 0).
</div>
```{r}
t.test(linec,linea,mu=0,conf.level = 0.95)
```
<div style="text-align: justify"> 
The null hypothesis that the means are equal (or that their difference is **0** ) is **not** trapped in the 95% confidence interval of the actual difference of the means [0.87,13.82], and because the p-value is smaller than the required 0.05 significance level, it can be **rejected** that they are equal, and assured _within the 5% s.l._ that they **differ**.
</div>
